---
title: "Week_8_Climate"
author: "Laura Wolf"
date: "5/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load Packages

```{r load_packages, include=FALSE}

# Load the libraries into this R session
library(raster)       #Main raster library with nearly all functions used in this analysis
library(rgdal)        #Spatial library - most functions used from rgdal are for vectors (shapefiles)
library(rasterVis)    #Useful for raster visualizations
library(maps)         #Has a database of maps. I use this to add a map to my raster to visualize land boundaries
library(rgeos)        #Need this library for topology operations on geometries
library(dplyr)        #NOT spatial - this is a data wrangling library
library(RColorBrewer) #Also not spatial - used to set the spectral color scheme
library(sp)
library(tmaptools)
library("tmap")

```

##Get some fun colors

```{r colors}

# rainbow color scheme
cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) 

#setting smaller margins for plotting
par(mar=c(2,2,1,1))

```

##Import Data

###Import species data

```{r import_species}

# Make sure the pathname to the data is correct
#Species
all_spp <- raster("~/github/conplan/ca_curr_sp_rich.tif")

all_spp #gives raster info


plot(all_spp,col=cols) #plots raster

#compare characteristics of species data to threats data below
#resolution is 0.5 degree squares, much lower than threats
#extent is different as well
#same projection as threats
#We need to crop and resample. Threats has larger extent, so we need change extent of threats. Species has lower resolution so we need to resample species based on threats. We will do this down below

```

###Import climate threats data

```{r import_threats}

# Change the path to the path to the lab data


#Threats
acid <- raster("~/github/conplan/impact_acid.tif")
uv <- raster("~/github/conplan/impact_uv.tif")
sst <- raster("~/github/conplan/impact_sst.tif")


#View characteristics of all_threats
acid
uv
sst

```

### Stack threats

```{r stack_threats}

cc_threat_stack <- stack(acid, uv, sst) #stacks threats so you can just call this instead of calling things individually for each threat
plot(cc_threat_stack,col=cols) #plots all climate threats next to each other

```

##Explore threat data

```{r explore_threat_data}

hist(cc_threat_stack)

```

Crazy Numbers! Let's reclassify them as NA and see what happens

###Reclassify Crazy Numbers

```{r reclass_cc_threats}

#get rid of crazy numbers, then reexamine histogram

acid_rcl <- reclassify(acid,rcl=c(-Inf,0,NA)) 
uv_rcl <- reclassify(uv,rcl=c(-Inf,0,NA))
sst_rcl <- reclassify(sst,rcl=c(-Inf,0,NA))
all_cc <- stack(acid_rcl, uv_rcl, sst_rcl)
hist(all_cc)
cellStats(all_cc,mean) 
cellStats(all_cc, sd) 
cellStats(all_cc, max)

```

That's better. Now let's do some cool stuff!

##Hotspots

###Function

```{r quantile_function}

#Makes a function to return a binary raster where the top X% of raster values are given a value of 1

reclassify_topx <- function(rast,quant=0.8) {
  topx <- quantile(rast,quant) #find the X% quantile of the raster values
  maxVal <- cellStats(rast,max) #find the maximum
  rcl <- c(-Inf,topx,0,
            topx,maxVal,1) # reclassify matrix (see help file for ?reclassify) to set values below X% quantile as 0, values above Xth quantile below max value as 1
  out <- reclassify(rast,rcl=rcl)
  return(out) # returns the new binary raster
}

```

###Apply function to climate threats

```{r reclass_cc_top_20}

#apply function from above to get top 20% of climate change threats and give them value of 1, give everything else value of 0
acid_top20 <- reclassify_topx(acid_rcl,quant=0.8)
uv_top20 <- reclassify_topx(uv_rcl,quant=0.8)
sst_top20 <- reclassify_topx(sst_rcl,quant=0.8)

```

###Combine climate threats

```{r merge_threats}

##FUNCTION TO ADD X + Y + Z
addXYZ <- function(x,y,z){
  return(x+y+z)
  }

combine_cc <- addXYZ(acid_top20,uv_top20,sst_top20) #adds top 20% of climate change threats together to get cumulative climate change threat layer.
plot(combine_cc, col=cols)

```

###Crop and Resample

####Crop

```{r crop}

threats_crop <- crop(combine_cc,all_spp) #Crop the threats layer to the same extent as species

```

####Resample

```{r resample}

spp_res <- resample(all_spp,combine_cc,method='ngb',progress='text') #resample species layer based on threats layer. the progress='text' argument prints out the progress of a longer-running function into the console, so you can see how the operation is going
# the method='ngb' argument specifies that we want to use a nearest neighbor algorithm to resample, instead of interpolation. interpolation would take average values.

```

```{r zeros}

#reclass zeros as NA
# we are OVERWRITING the original spp_res object.
# This is okay in this instance since we won't be using the old version, but
# often it is better to assign any output of a function to a new variable or object

spp_res <- reclassify(spp_res,rcl=c(-Inf,0,NA)) #default is inclusive
hist(spp_res,main="Species Raster Values, Zeroes Removed") # did the function do what we were hoping? Compare histogram to the one above. Large column at zero has gone to zero!

```

###Make species data binary

```{r binary_species}

spp_binary <- reclassify_topx(spp_res, quant = 0.8) #applies function to species data to find the top 20% and set that to 1, with everything else set to 0

```

###Combine species and threats layers to get hotspots

```{r overlay, eval=FALSE, include=FALSE}

##FUNCTION TO ADD X + Y
addXY <- function(x,y){
  return(x+y)
  }

hotspots <- overlay(spp_binary, climate_top20, fun=addXY) #adds top 20% of species to top 20% of climate threats

# color breakpoints. We need three colors now! (cell values of 0,1,or 2)
brks_hotspots <- seq(0,3,length.out=4) 
hotspot_cols <- c("white","white","red") #firebrick shows where top20% of threats overlap with top 20% of species richness. Not that many places...

# plot the hotspots!
plot(hotspots,col=hotspot_cols,legend=F,main="Hotspots");map('world',fill=T,add=T,col='gray80')

```


##Mean Threat

```{r mean_threat}

mean_threat <- overlay(acid_rcl, uv_rcl, sst_rcl, fun=function(x,y,z) {return((x+y+z)/3)}) #makes a new raster with the average threat value in cell
plot(mean_threat, col=cols)

```


```{r crop_mean}

mean_crop <- crop(mean_threat,spp_binary) #crop mean to same extent at top 20% of species

```

###Combine mean threats with top 20% of hotspots

```{r overlay_mean}



combine_mean_spp <- overlay(spp_binary, mean_crop, fun=addXY) #adds the mean threat value for each cell to the binary species values of the top 20% of species richness. This gives values from 0 ~ 1.8

combine_mean_spp_1 <- combine_mean_spp - 1 #subtracts 1 from the above raster to get values below 1

hotspots_mean <- reclassify(combine_mean_spp_1,rcl=c(-Inf,0,NA)) #assigns negative values as NA



# plot the hotspots!
plot(hotspots_mean,col=cols,legend=T,main="Hotspots");map('world',fill=T,add=T,col='gray80')

```

##MPAs

###Import data

```{r import_mpa}

#This is how you import shapefiles. See basemap example below for better, more specific method. The "." implies that the file is in the working folder. Layer is the specific layer name. 
mpa <- readOGR(dsn = ".", layer = "MPA_CA_Existing_160301")
sanctuary <- readOGR(dsn = ".", layer = "National Marine Sanctuaries")

# Transforms mpa layer and sactuary layer into same projection as mean threat layer
mpa_transform <- spTransform(mpa, crs(mean_threat))
sanctuary_transform <- spTransform(sanctuary, crs(mean_threat))


```

###Look at Data

```{r mpa_plot}

plot(mpa_transform);map('world',fill=T,add=T,col='gray80')
plot(sanctuary_transform, col=cols,main="National Marine Sanctuaries");map('world', fill=T,add=T,col='gray80')

```

### Crop 

```{r crop_means}

mean_crop_sanc <- crop(mean_threat,sanctuary_transform) #crops mean threat at extent of sanctuaries
mean_crop_mpa <- crop(mean_threat, mpa_transform) #crops mean threat at extent of mpas
hotspot_crop_ca <- crop(hotspots_mean, extent(-126,-117, 33, 42)) #crops hotspots of mean threats and top 20% of species richness at extent of California coastline
mean_crop_ca <- crop(mean_threat, extent(-126,-117, 33, 42)) #crops mean threats at extent of California coastline

plot(hotspot_crop_ca, col=cols,legend=T,main="Hotspots");map('world',fill=T,add=T,col='gray80')

```
###Import Basemap for Figure

```{r import_basemap}

#("C:\Users\Brad\Documents\github\conplan\cb_2017_us_nation_5m")

#Better way of importing shapefiles
dir_spatial   <- '/Users/Brad/Documents/github/conplan/cb_2017_us_nation_5m' #Location of file
layer_us <- 'cb_2017_us_nation_5m' #name of file

us_base <- readOGR(dsn = dir_spatial, layer = layer_us, stringsAsFactors = FALSE)

us_transform <- spTransform(us_base, crs(mean_threat)) #transforms projection to same as mean_threat

```

###Map Results!

```{r tmap}

tm_mean <- tm_shape(mean_crop_ca)+ #adds mean threat to map
  tm_raster("layer", #displays as raster by name of element in raster
            palette = cols, #color palette
            title = "Mean Threat", #title that shows up in legend
            breaks=c(0.5, 0.6, 0.7, 0.8, 0.9), #breaks that show up in legend
            legend.hist = TRUE)+ #shows histogram
tm_shape(us_base)+ #adds basemap
  tm_polygons()+ #displays as polygon
tm_shape(sanctuary_transform)+ #adds sanctuaries
  tm_borders(NA)+ #only displays borders
tm_style_white() + #color scheme of whole map (white backgroud, grey fill of US)
  tm_legend(text.size=1,
    title.size=1.3,
    position = c(0.68,0.48), 
    bg.alpha=.2, 
    frame="gray50", 
    height=.4, 
    hist.width=.28,
    hist.height=.2, 
    hist.bg.color="gray60", 
    hist.bg.alpha=.5)+
tm_compass(type = "arrow", #style of compass
           position = c("right", "top"),
           color.light = "black", #makes whole thing black
           size = 1)+
tm_scale_bar(position = c(0.55,0.9))

tm_hotspot <- tm_shape(hotspot_crop_ca)+
  tm_raster("layer",
            palette = cols,
            title = "Mean Threat",
            breaks=c(0.5, 0.6, 0.7, 0.8, 0.9),
            legend.hist = TRUE)+
tm_shape(us_base)+
  tm_polygons()+
tm_shape(sanctuary_transform)+
  tm_borders(NA)+
tm_style_white() +
  tm_legend(text.size=1,
    title.size=1.3,
    position = c(0.68,0.48), 
    bg.alpha=.2, 
    frame="gray50", 
    height=.4, 
    hist.width=.28,
    hist.height=.2, 
    hist.bg.color="gray60", 
    hist.bg.alpha=.5)+
tm_compass(type = "arrow",
           position = c("right", "top"),
           color.light = "black",
           size = 1)+
tm_scale_bar(position = c(0.55,0.9))

combined_map <- tmap_arrange(tm_mean, tm_hotspot, asp = NA) #displays the two maps next to each other


```

```{r save_map}

save_tmap(tm_hotspot, "wk_8_hotspots.png")
save_tmap(tm_mean, "wk_8_mean.png")

```
####FINISH HOT SPOTS MAP

```{r}

#Combine three climate threats, then find the top 20% highest threat areas

climate_threats <- addXYZ(acid_rcl,uv_rcl,sst_rcl)
climate_top20 <- reclassify_topx(climate_threats,quant=0.8)


#add top 20% of species to top 20% of climate
hotspots <- overlay(spp_binary, climate_top20, fun=addXY)

brks_hotspots <- seq(0,3,length.out=4) 
hotspot_cols <- c("white","white","red") #firebrick shows where top20% of threats overlap with top 20% of species richness. Not that many places...

# plot the hotspots!
plot(hotspots,col=hotspot_cols,legend=F,main="Hotspots");map('world',fill=T,add=T,col='gray80')

```


```{r}
hs_crop_sanc <- crop(hotspots,sanctuary_transform) #crops hotspots at extent of sanctuaries
hs_crop_mpa <- crop(hotspots, mpa_transform) #crops hotspots at extent of mpas
hotspot_crop_wa <- crop(hotspots, extent(-130,-122, 40, 50)) #crops hotspots of mean threats and top 20% of species richness at extent of Washington & Oregon coastline
threat_crop_wa <- crop(climate_top20, extent(-130,-122, 40, 50)) #crops mean threats at extent of Washington & Oregon coastline

plot(hotspot_crop_wa, col=hotspot_cols,legend=T,main="Hotspots");map('world',fill=T,add=T,col='gray80')

```


```{r}

tm_mean <- tm_shape(hotspot_crop_wa)+ #adds mean threat to map
  tm_raster("layer", #displays as raster by name of element in raster
            palette = hotspot_cols) #color palette
tm_shape(us_base)+ #adds basemap
  tm_polygons()+ #displays as polygon
tm_shape(sanctuary_transform)+ #adds sanctuaries
  tm_borders(NA)+ #only displays borders
tm_style_white() + #color scheme of whole map (white backgroud, grey fill of US)
  tm_legend(text.size=1,
    title.size=1.3,
    position = c(0.68,0.48), 
    bg.alpha=.2, 
    frame="gray50", 
    height=.4, 
    hist.width=.28,
    hist.height=.2, 
    hist.bg.color="gray60", 
    hist.bg.alpha=.5)+
tm_compass(type = "arrow", #style of compass
           position = c("right", "top"),
           color.light = "black", #makes whole thing black
           size = 1)+
tm_scale_bar(position = c(0.55,0.9))

tm_hotspot <- tm_shape(hotspot_crop_ca)+
  tm_raster("layer",
            palette = cols,
            title = "Mean Threat",
            breaks=c(0.5, 0.6, 0.7, 0.8, 0.9),
            legend.hist = TRUE)+
tm_shape(us_base)+
  tm_polygons()+
tm_shape(sanctuary_transform)+
  tm_borders(NA)+
tm_style_white() +
  tm_legend(text.size=1,
    title.size=1.3,
    position = c(0.68,0.48), 
    bg.alpha=.2, 
    frame="gray50", 
    height=.4, 
    hist.width=.28,
    hist.height=.2, 
    hist.bg.color="gray60", 
    hist.bg.alpha=.5)+
tm_compass(type = "arrow",
           position = c("right", "top"),
           color.light = "black",
           size = 1)+
tm_scale_bar(position = c(0.55,0.9))

combined_map <- tmap_arrange(tm_mean, tm_hotspot, asp = NA) #displays the two maps next to each other

```


